Financialisation of Commodity Markets

### 2025-07-04 docs/paper.qmd
Update paper
Replace old tables with new; update body of text and captions accordingly

@tbl-stats-combined displays descriptive statistics for the commodity series considered in the study; the results show a clear pattern of increase in average return during financialisation with an equally weighted portfolio of the US traded assets showing a mean returns of 15.9% significant at the 5% level vs. 6.9% with a 10% significance in the pre-financialisation period. The increase is even more pronounced for an equally weighted portfolio of GB traded assets (metals), 2.47% (not significant) to 21.8% (significant at the 10% level), which sets out the global nature of the phenomenon. Most commodities show a higher mean return while five---Lumber (XCME), Natural gas (XNYM), Palladium (XNYM), Platinum (XNYM) and Nickel-primary (XLME)---show a lower figure with five of them significant---Crude oil-WTI (XNYM), Heating oil (XNYM), Copper (XCEC), Copper (XLME) and Tin-refined (XLME)---against none pre-financialisation. Sugar-#11 wins the race with a mean return increasing from -4.6% to 24.9% followed by copper with sizeable increases for both markets: 1.5% to 28.2% (XCEC) and 0.5% to 28.6% (XLME).  

@@

The cross-sectional nature of financialisation naturally leads us to extend the individual hedging pressure paradigm by incorporating aggregate CHP (**CHP**), a market wide measure of hedging pressure. The traditional Keynesian hedging pressure paradigm postulates a negative relationship between the return on an individual commodity futures and its hedging pressure. The extended Keynesian hedging pressure paradigm, in the context of financialisation, should also imply that returns for individual commodities should be high in periods where **CHP** is low and vice et versa. @tbl-stats-combined further shows the mean returns and volatilities for high (**contango**) and low (**backwardation**) **CHP** regimes for each time period while @tbl-regime-diference-tests shows the results of difference tests for returns mean and variance between the two regimes.  

@@

Financialisation was a global phenomenon, through the channel of index investment it simultaneously influenced the entire spectrum of liquid commodities, eventually spurring the transformation of commodity markets from a collection of idiosyncratic markets into a more cohesive collection akin to an asset class. In this context, interconnectedness emerges as a crucial aspect that warrants thorough examination.   
We strive to address this issue by first examining the average returns pairwise correlations between commodities across the entire cross-section of the commodity assets considered in the study as well as by country, sector and sub-sector. @tbl-correlations-inner-periods shows the results for the four periods of interest with the results reported for the entire period as well as for periods of **backwardation** and **contango**. For the entire sample the average pairwise correlation jumps from 0.07 in the first period to 0.16 over financialisation. This finding indicates higher interdependence within the full cross-section of commodities starting with the financialisation period, consistent with the empirical findings in @fryMcKibbin_evolution_2023 and @mayer_financialization_2017 as well as with the theoretical model of @basak_model_2016. The sector analysis sheds light on the driving force behind these changes. The US metals see the sharpest increase in pairwise correlations going from 0.22 over the past period to 0.56 over financialisation while the three other sectors see a much smaller percentage increase during financialisation relative to the past period. These results thus suggest that the US metals sector was the driving force behind increased average pairwise correlations during financialisation, a new observation to the best of our knowledge. The pattern of correlations is very similar across phases of **backwardation** and **contango** for all the sectors and sub-sectors over the past and financialisation periods. In this context, the ontological consequences of financialisation are better observed from an investment perspective as illustrated in @tbl-correlations-cross-periods where we observe average pairwise correlation coefficients between equally weighted portfolios of US traded commodity assets grouped by sub-sector^[A more detailed breakdown of country, sector and subsector average pairwise correlations is provided in the attached [Internet Appendix](http://bautheac.duckdns.org:3838/bautheac/revision-jfm/).]. Although modest in magnitude, these exhibit a telling pattern, mirroring that shown by the individual commodity returns in @tbl-stats-combined, with a substantially higher average coefficient over **backwardation** phases in the first period (0.12 vs. 0.03) and the pattern reversed over financilisation with an average coefficient substantially higher in **contango** (0.12 vs. 0.02).  
We break down @tbl-correlations-inner-periods's average correlations by year in @tbl-correlations-inner-years. For the full sample of assets, the average pairwise correlations remain in the 0.06–0.08 range from 1998 through 2003, before suddenly increasing to 0.12 from 2003 to 2004. This suggests 2004 as a distinct break point for the onset of financialisation, as noted in several studies [@baker_financialization_2021; @tang_index_2012]. We further observe that the US metals appear to have played a major role in this as the increase seems driven by the sector which average pairwise correlation more than doubles (0.49 vs 0.22) over the period.

@@

The corresponding pricing performance displays a similar pattern as shown in @tbl-regressions-factors where for each period the corresponding $R^{2}$ values are independently averaged over the entire period as well as over periods of **contango** and **backwardation**. The average $R^{2}$ for the market factor across the entire cross-section of commodity assets, including the GB traded base metals, increases to around 20% during financialisation compared to around 10% over the previous period with the rise driven by the metals sector at large. The sector average for the US traded assets jumps to 31% during financialisation from 7.5% over the past period while for the GB traded assets the corresponding figure increases from less than 2%, or virtually no explanatory power for the (US based) market factor, to almost 12% over the financialisation period. The results for our other risk factor mimicking portfolios show a similar pattern and provide a deeper insights into co-movement dynamics in this context. All three factors---CHP, open interest and term structure---performed poorly in the first period, indicating that the commodities with extreme factor attributes (high or low CHP for example) were uncorrelated with the rest of the commodity cross-section, and the overall performance consistent across periods of **contango** and **backwardation**. The onset of financialisation leads to a substantial improvement in the performance of the CHP factor with its average $R^{2}$, albeit still small at 5.4%, almost tripling from the past to the financialisation period in contrast to the other factors that show weaker dynamics with a slight decrease and a much smaller increase for the open interest and term structure factors respectively. Here again the increase in pricing performance for the CHP factor is driven by the metals sector, particularly the precious sub-sector with the average $R^{2}$ for the group jumping to almost 26% during financialisation from 1.2% over the past period. The corresponding figures of 22% and 1.1% for the entire US metals sector stand in sharp contrast with the agricultural sector, which exhibits no notable change, and the energy sector, where pricing performance drops substantially from 9.2% down to 1.6%. The improvement in the CHP factor's pricing performance remarkably comes predominantly from the long leg (low CHP assets: Keynesian "normal backwardation") of the factor^[Results available from the authors upon request.] with the corresponding mimicking portfolio's performance itself driven by the metals, the precious sub-sector in particular. Interestingly, although the term structure factor itself only sees a very modest increase, its short leg (low roll-yield assets: Working's backwardation) shows a sizeable rise in pricing performance, yet again driven the metals sector. The pattern of results for both factors extends to the GB traded assets and suggest a convergence of Keneysian and Working's term structure perspective over the financialisation period.   

@@

The crisis period (late 2008 to mid-2013) shows a contrasting pattern of results. Returns fall sharply relative to the financialisation period (@tbl-stats-combined) with eighteen US commodities along with all six GB traded metals showing lower mean returns. Returns for equally weighted portfolios of both sets of assets fall down to 9.4% and 5.3% from 15.9% and 21.8% respectively. Returns are mostly higher during phases of **backwardation** relative to **contango** with fifteen US assets and five of the GB traded metals showing higher mean returns in **backwardation** and the pattern extending to equally weighted portfolios of the two sets of assets. Volatility is moderately higher for both sets with fifteen US commodities and three GB metals showing higher figures and interestingly eminently higher over phases of **contango**; out of the US complex only Sugar-#11 (IFUS) and none of the GB traded metals show higher volatility figures over **backwardation** phases. The symmetry between the dynamics of CHP and **CHP**'s explanatory power on individual commodity returns breaks (@tbl-time-reg-US-CHP) with both now seemingly converging; the US cross-sectional average $R^{2}$ at the individual level bouncing back up (22.5% vs 20.8%) while that at the aggregate level rises further (14.7% vs 7.1%).

@@

The returns pull-back continues into the post-crisis period (@tbl-stats-combined) with twenty-two US and four GB traded assets showing lower mean returns relative to the crisis while the same number of US commodities and all GB metals show lower figures relative to the financialisation period; mean returns for equally weighted portfolios of US and GB assets drop down to -0.2% and 2.7% respectively. Interestingly volatility also recedes substantially with twenty US assets and all GB metals showing lower volatility figures relative to both the crisis and financialisation periods suggesting that investment outflows from the commodity futures complex have had a lower impact on volatility than the corresponding inflows witnessed over the financialisation period. The figures for the equally weighted portfolios of US and GB assets drop to 9.9% and 16.2% respectively, very close to pre-financialisation levels.  
The pattern of reversion toward the pre-financialisation Keynesian paradigm initiated during the crisis continues stronger into the post-crisis period with mean returns higher during phases of **backwardation** relative to **contango**, more so than over the crisis period with seventeen US commodities showing higher mean returns over **backwardation** phases. The difference (@tbl-regime-diference-tests) is significant at the 5% level for Lumber (XCME) and at the 10% level for Palladium (XNYM) while no difference is significant for the assets showing higher figures over phases of **contango**. The pattern is more pronounced for the GB traded subset with all six assets showing higher mean returns in **backwardation** with the difference significant at the 1% level for Nickel-primary (XLME), 5% level for Aluminium (XLME) and 10% level for Zinc (XLME). It also extends to equally weighted portfolios of US and GB traded assets that both show higher mean returns during **backwardation** with the difference significant at the 5% level for the GB portfolio and the figure close to the 10% level significance threshold for the portfolio of US assets. Volatilities are overall higher in **contango** although less so than over the previous period with seventeen US commodities showing higher figures over **contango** against twenty-three in the crisis period. The pattern extends to three of the six GB traded metals with the difference significant at the 1% level for Copper (XLME) and Tin (XLME) and at the 5% level for Zinc (XLME) while differences are not significant for Aluminium (XLME), Lead (XLME) and Nickel-primary (XLME) that exhibit higher volatility over **backwardation** phases. Equally weighted portfolios of US and GB traded assets on the other hand both show higher volatility in **contango** with the difference significant at the 1% level for both. Over the financialisation period in contrast volatilities were predominantly higher in **backwardation** for US traded assets with fourteen commodities showing higher volatility figures in these phases; the post-crisis pattern of results for volatilities thus likewise shows signs of reversion to pre-financialisation standards. After showing signs of convergence in the crisis period the dynamics of the explanatory power of CHP and **CHP** part ways (@tbl-time-reg-US-CHP), seemingly reverting to the Keynesian paradigm observed pre-financialisation, further suggesting a return to legacy commodity specific dynamics. The recovery initiated in the crisis period for that of CHP is confirmed with a stronger increase, 27.2% vs. 22.5%, strongest for the metals sector, 37.3% vs. 23.7%, while the energy sector shows the opposite pattern: 11% vs. 16%. In contrast, after increasing stronger in the crisis, that for **CHP** falls overall post-crisis, 7.5% vs 14.7%, with the decrease strongest for the energy sector in a relative sense: 2.5% vs. 14.8%.

@@

```{r}
#| label: tbl-stats-combined
#| tbl-cap: "This table shows mean returns and corresponding volatility for the twenty four individual US commodities and the six GB metals considered in the study as well as for two equally weighted portfolios formed from the US commodities and the GB metals respectively across the four periods of interest (past: 1997-2003; financialisation: 2004-2008; crisis: 2008-2013; post-crisis: 2013-2018). for each period the figures are shown independently for the entire period as well as for phases of aggregate backwardation (aggregate CHP $\\leq$ period median) and aggregate contango (aggregate CHP > period median). Mean values significant at the 1%, 5% and 10% level are marked with \\\\*\\\\*\\\\*, \\\\*\\\\* and \\\\* respectively. Aggregate CHP construction and corresponding regime definitions are discussed in \\@sec-methods while the results are discussed in \\@sec-results."

table_name <- "stats - combined"
columns_to_collapse_rows_for <- c("asset", "regime")
colnames <- c("asset", "regime", "estimate", periods)
align <- make_series_of_repeated_chars(list(l = 3L, r = 4L))
longtable <- TRUE
bold_columns <- 1L:3L
coloured_columns <- 4L:7L

@@

```{r}
#| label: tbl-regime-diference-tests
#| tbl-cap: "This table shows the results for tests of difference in mean (t-test) and variance (F-test) between phases of aggregate backwardation and contango for the twenty four individual US commodities and the six GB metals considered in the study as well as for two equally weighted portfolios formed from the US commodities and the GB metals respectively across the four periods of interest (past: 1997-2003; financialisation: 2004-2008; crisis: 2008-2013; post-crisis: 2013-2018). 1%, 5% and 10% p-value levels are marked with \\\\*\\\\*\\\\*, \\\\*\\\\* and \\\\* respectively in the significance column. Aggregate CHP construction and corresponding regime definitions are discussed in \\@sec-methods while the results are discussed in \\@sec-results."

table_name <- "regime difference tests"
columns_to_collapse_rows_for <- c("asset", "period")
colnames <- c("asset", "period", "moment", "dominant regime", "p-value", "significance")
align <- make_series_of_repeated_chars(list(l = 4L, r = 2L))
longtable <- TRUE
bold_columns <- 1L:4L 
coloured_columns <- 5L:6L

```{r}
#| label: tbl-time-reg-US-CHP
#| tbl-cap: "This table shows the average time series $R^{2}$ for models where the returns series for the twenty four individual US commodities are independently regressed against relative change in their own commercial hedging pressure (CHP) series as well as against that in aggregate CHP, a market wide measure of CHP which calculation method is discussed in \\@sec-methods. Averages are calculated independently across the whole cross-section of US traded commodity assets considered in the study as well as across commodity sectors and sub-sectors for the four periods of interest (past: 1997-2003; financialisation: 2004-2008; crisis: 2008-2013; post-crisis: 2013-2018). The results are discussed in \\@sec-results while \\@sec-taxonomy details the taxonomy for the commodity assets considered."

@@

```{r}
#| label: tbl-correlations-inner-periods
#| tbl-cap: "This table shows the average return pairwise correlation coefficients calculated independently across the whole cross-section of commodity assets considered in the study as well as within countries and commodity sectors and sub-sectors for the four periods of interest (past: 1997-2003; financialisation: 2004-2008; crisis: 2008-2013; post-crisis: 2013-2018). For each period, the results are calculated independently over the entire period as well as over phases of aggregate backwardation and contango. Aggregate CHP construction and corresponding regime definitions are discussed in \\@sec-methods while the results are discussed in \\@sec-results and the taxonomy for the commodity assets considered is detailed in \\@sec-taxonomy."

@@

```{r}
#| label: tbl-correlations-cross-periods
#| tbl-cap: "This table shows the average return pairwise correlation coefficients between equally weighted US sub-sector portfolios (agriculturals: gains, livestock, softs; energy: gas, petroleum; metals: base, precious). The results are independently calculated for the four periods of interest (past: 1997-2003; financialisation: 2004-2008; crisis: 2008-2013; post-crisis: 2013-2018). For each period, the results are further calculated independently over the entire period as well as over phases of aggregate backwardation and contango. Aggregate CHP construction and corresponding regime definitions are discussed in \\@sec-methods while the results are discussed in \\@sec-results and the taxonomy for the commodity assets considered is detailed in \\@sec-taxonomy."

table_name <- "correlations - cross - US - periods - subsectors"
columns_to_collapse_rows_for <- c("regime")
colnames <- c("regime", periods)
align <- make_series_of_repeated_chars(list(l = 1L, r = 4L))
longtable <- TRUE
bold_columns <- 1L
coloured_columns <- 2L:5L

@@

```{r}
#| label: tbl-correlations-inner-years
#| tbl-cap: "While \\@tbl-correlations-inner-periods shows results calculated independently for the four periods of interest (past: 1997-2003; financialisation: 2004-2008; crisis: 2008-2013; post-crisis: 2013-2018), this stable shows the results of the same analysis calculated by year over the entire sample period (1997-2018). The results are discussed in \\@sec-results while \\@sec-taxonomy details the taxonomy for the commodity assets considered."

@@

```{r}
#| label: tbl-regressions-index
#| tbl-cap: "This table shows the average time series regression coefficients for models where the returns series for all the commodity assets considered in the study are independently regressed against the return series for our market factor, an equally weighted portfolio of the 24 US commodities considered in the study. The results are calculated independently across countries as well as commodity sectors and sub-sectors for the four periods of interest (past: 1997-2003; financialisation: 2004-2008; crisis: 2008-2013; post-crisis: 2013-2018). The market factor construction is discussed in \\@sec-methods while the results are discussed in \\@sec-results and the taxonomy for the commodity assets considered is detailed in \\@sec-taxonomy."

@@

```{r}
#| label: tbl-regressions-factors
#| tbl-cap: "This table shows the average time series regression $R^{2}$ for models where the returns series for all the commodity assets considered in the study are independently regressed against the return series for our mimicking portfolios for risk factors in returns, including the market, CHP, open interest and term structure factors. The results are calculated independently across countries and commodity sectors and sub-sectors for the four periods of interest (past: 1997-2003; financialisation: 2004-2008; crisis: 2008-2013; post-crisis: 2013-2018); for each period the results are calculated independently over the entire period as well as over phases of aggregate backwardation and contango. Factors as well as aggregate CHP construction and corresponding regime definitions are discussed in \\@sec-methods while the results are discussed in \\@sec-results and the taxonomy for the commodity assets considered is detailed in \\@sec-taxonomy."

table_name <- "regressions - all returns ~ factors"
columns_to_collapse_rows_for <- c("country", "sector", "subsector", "factor")
colnames <- c(sectors, "factor", "regime", periods)
align <- make_series_of_repeated_chars(list(l = 5L, r = 4L))
longtable <- TRUE
bold_columns <- 1L:9L
coloured_columns <- 6L:9L

@@

```{r}
#| label: tbl-assets-taxonomy
#| tbl-cap: "This table shows the detailed taxonomy for the thirty commodity assets considered in the study. The assets are classified by the country of residence of the exchange where they are traded as well as by sector and sub-sector according to the nature of the asset."

>>>>

@tbl-stats-regime-tests displays descriptive statistics for the commodity series considered in the study; the results show a clear pattern of increase in average return during financialisation with an equally weighted portfolio of the US traded assets showing a mean returns of 15.9% significant at the 5% level vs. 6.9% with a 10% significance in the pre-financialisation period. The increase is even more pronounced for an equally weighted portfolio of GB traded assets (metals), 2.47% (not significant) to 21.8% (significant at the 10% level), which sets out the global nature of the phenomenon. Most commodities show a higher mean return while five---Lumber (XCME), Natural gas (XNYM), Palladium (XNYM), Platinum (XNYM) and Nickel-primary (XLME)---show a lower figure with five of them significant---Crude oil-WTI (XNYM), Heating oil (XNYM), Copper (XCEC), Copper (XLME) and Tin-refined (XLME)---against none pre-financialisation. Sugar-#11 wins the race with a mean return increasing from -4.6% to 24.9% followed by copper with sizeable increases for both markets: 1.5% to 28.2% (XCEC) and 0.5% to 28.6% (XLME).  

@@

The cross-sectional nature of financialisation naturally leads us to extend the individual hedging pressure paradigm by incorporating aggregate CHP (**CHP**), a market wide measure of hedging pressure. The traditional Keynesian hedging pressure paradigm postulates a negative relationship between the return on an individual commodity futures and its hedging pressure. The extended Keynesian hedging pressure paradigm, in the context of financialisation, should also imply that returns for individual commodities should be high in periods where **CHP** is low and vice et versa. @tbl-stats-regime-tests further reports the dominating regime for each asset-period-moment combination along with the corresponding results of difference tests for returns mean and variance between the two regimes^[The estimate values per regime as well as the p-values for the corresponding difference tests are available in the attached [Internet Appendix](http://bautheac.duckdns.org:3838/bautheac/co-movement/).].  

@@

Financialisation was a global phenomenon, through the channel of index investment it simultaneously influenced the entire spectrum of liquid commodities, eventually spurring the transformation of commodity markets from a collection of idiosyncratic markets into a more cohesive collection akin to an asset class. In this context, interconnectedness emerges as a crucial aspect that warrants thorough examination.   
We strive to address this issue by first examining the average returns pairwise correlations between commodities across the entire cross-section of the commodity assets considered in the study as well as by country, sector and sub-sector. @tbl-correlations-inner-periods shows the results for the four periods of interest with the results reported for the entire period as well as for periods of **backwardation** and **contango**. For the entire sample the average pairwise correlation jumps from 0.07 in the first period to 0.16 over financialisation. This finding indicates higher interdependence within the full cross-section of commodities starting with the financialisation period, consistent with the empirical findings in @fryMcKibbin_evolution_2023 and @mayer_financialization_2017 as well as with the theoretical model of @basak_model_2016. The sector analysis sheds light on the driving force behind these changes. The US metals see the sharpest increase in pairwise correlations going from 0.22 over the past period to 0.56 over financialisation while the three other sectors see a much smaller percentage increase during financialisation relative to the past period. These results thus suggest that the US metals sector was the driving force behind increased average pairwise correlations during financialisation, a new observation to the best of our knowledge. The pattern of correlations is very similar across phases of **backwardation** and **contango** for all the sectors and sub-sectors over the past and financialisation periods. In this context, the ontological consequences of financialisation are better observed from an investment perspective as illustrated in @tbl-correlations-cross-periods where we observe average pairwise correlation coefficients between equally weighted portfolios of US traded commodity assets grouped by sub-sector^[A more detailed breakdown of country, sector and subsector average pairwise correlations is provided in the attached [Internet Appendix](http://bautheac.duckdns.org:3838/bautheac/co-movement/).]. Although modest in magnitude, these exhibit a telling pattern, mirroring that shown by the individual commodity returns in @tbl-stats-regime-tests, with a substantially higher average coefficient over **backwardation** phases in the first period (0.12 vs. 0.03) and the pattern reversed over financilisation with an average coefficient substantially higher in **contango** (0.12 vs. 0.02).  
We break down @tbl-correlations-inner-periods's average correlations by year in @tbl-correlations-inner-years. For the full sample of assets, the average pairwise correlations remain in the 0.06–0.08 range from 1998 through 2003, before suddenly increasing to 0.12 from 2003 to 2004. This suggests 2004 as a distinct break point for the onset of financialisation, as noted in several studies [@baker_financialization_2021; @tang_index_2012]. We further observe that the US metals appear to have played a major role in this as the increase seems driven by the sector which average pairwise correlation more than doubles (0.49 vs 0.22) over the period.

@@

The corresponding pricing performance displays a similar pattern as shown in @tbl-regressions-factors where for each period the corresponding $R^{2}$ values are independently averaged over the entire period as well as over periods of **contango** and **backwardation**. The average $R^{2}$ for the market factor across the entire cross-section of commodity assets, including the GB traded base metals, increases to around 20% during financialisation compared to around 10% over the previous period with the rise driven by the metals sector at large. The sector average for the US traded assets jumps to 31% during financialisation from 7.5% over the past period while for the GB traded assets the corresponding figure increases from less than 2%, or virtually no explanatory power for the (US based) market factor, to almost 12% over the financialisation period. The results for our other risk factor mimicking portfolios show a similar pattern and provide a deeper insights into co-movement dynamics in this context. All three factors---CHP, open interest and term structure---performed poorly in the first period, indicating that the commodities with extreme factor attributes (high or low CHP for example) were uncorrelated with the rest of the commodity cross-section, and the overall performance consistent across periods of **contango** and **backwardation**^[A breakdown of the performance by regime is provided in the attached [Internet Appendix](http://bautheac.duckdns.org:3838/bautheac/co-movement/).]. The onset of financialisation leads to a substantial improvement in the performance of the CHP factor with its average $R^{2}$, albeit still small at 5.4%, almost tripling from the past to the financialisation period in contrast to the other factors that show weaker dynamics with a slight decrease and a much smaller increase for the open interest and term structure factors respectively. Here again the increase in pricing performance for the CHP factor is driven by the metals sector, particularly the precious sub-sector with the average $R^{2}$ for the group jumping to almost 26% during financialisation from 1.2% over the past period. The corresponding figures of 22% and 1.1% for the entire US metals sector stand in sharp contrast with the agricultural sector, which exhibits no notable change, and the energy sector, where pricing performance drops substantially from 9.2% down to 1.6%. The improvement in the CHP factor's pricing performance remarkably comes predominantly from the long leg (low CHP assets: Keynesian "normal backwardation") of the factor^[A breakdown of the performance by factor leg is provided in the attached [Internet Appendix](http://bautheac.duckdns.org:3838/bautheac/co-movement/).] with the corresponding mimicking portfolio's performance itself driven by the metals, the precious sub-sector in particular. Interestingly, although the term structure factor itself only sees a very modest increase, its short leg (low roll-yield assets: Working's backwardation) shows a sizeable rise in pricing performance, yet again driven the metals sector. The pattern of results for both factors extends to the GB traded assets and suggest a convergence of Keneysian and Working's term structure perspective over the financialisation period.   

@@

The crisis period (late 2008 to mid-2013) shows a contrasting pattern of results. Returns fall sharply relative to the financialisation period (@tbl-stats-regime-tests) with eighteen US commodities along with all six GB traded metals showing lower mean returns. Returns for equally weighted portfolios of both sets of assets fall down to 9.4% and 5.3% from 15.9% and 21.8% respectively. Returns are mostly higher during phases of **backwardation** relative to **contango** with fifteen US assets and five of the GB traded metals showing higher mean returns in **backwardation** and the pattern extending to equally weighted portfolios of the two sets of assets. Volatility is moderately higher for both sets with fifteen US commodities and three GB metals showing higher figures and interestingly eminently higher over phases of **contango**; out of the US complex only Sugar-#11 (IFUS) and none of the GB traded metals show higher volatility figures over **backwardation** phases. The symmetry between the dynamics of CHP and **CHP**'s explanatory power on individual commodity returns breaks (@tbl-time-reg-US-CHP) with both now seemingly converging; the US cross-sectional average $R^{2}$ at the individual level bouncing back up (22.5% vs 20.8%) while that at the aggregate level rises further (14.7% vs 7.1%).

@@

The returns pull-back continues into the post-crisis period (@tbl-stats-regime-tests) with twenty-two US and four GB traded assets showing lower mean returns relative to the crisis while the same number of US commodities and all GB metals show lower figures relative to the financialisation period; mean returns for equally weighted portfolios of US and GB assets drop down to -0.2% and 2.7% respectively. Interestingly volatility also recedes substantially with twenty US assets and all GB metals showing lower volatility figures relative to both the crisis and financialisation periods suggesting that investment outflows from the commodity futures complex have had a lower impact on volatility than the corresponding inflows witnessed over the financialisation period. The figures for the equally weighted portfolios of US and GB assets drop to 9.9% and 16.2% respectively, very close to pre-financialisation levels.  
The pattern of reversion toward the pre-financialisation Keynesian paradigm initiated during the crisis continues stronger into the post-crisis period with mean returns higher during phases of **backwardation** relative to **contango**, more so than over the crisis period with seventeen US commodities showing higher mean returns over **backwardation** phases. The difference (@tbl-stats-regime-tests) is significant at the 5% level for Lumber (XCME) and at the 10% level for Palladium (XNYM) while no difference is significant for the assets showing higher figures over phases of **contango**. The pattern is more pronounced for the GB traded subset with all six assets showing higher mean returns in **backwardation** with the difference significant at the 1% level for Nickel-primary (XLME), 5% level for Aluminium (XLME) and 10% level for Zinc (XLME). It also extends to equally weighted portfolios of US and GB traded assets that both show higher mean returns during **backwardation** with the difference significant at the 5% level for the GB portfolio and the figure close to the 10% level significance threshold for the portfolio of US assets. Volatilities are overall higher in **contango** although less so than over the previous period with seventeen US commodities showing higher figures over **contango** against twenty-three in the crisis period. The pattern extends to three of the six GB traded metals with the difference significant at the 1% level for Copper (XLME) and Tin (XLME) and at the 5% level for Zinc (XLME) while differences are not significant for Aluminium (XLME), Lead (XLME) and Nickel-primary (XLME) that exhibit higher volatility over **backwardation** phases. Equally weighted portfolios of US and GB traded assets on the other hand both show higher volatility in **contango** with the difference significant at the 1% level for both. Over the financialisation period in contrast volatilities were predominantly higher in **backwardation** for US traded assets with fourteen commodities showing higher volatility figures in these phases; the post-crisis pattern of results for volatilities thus likewise shows signs of reversion to pre-financialisation standards. After showing signs of convergence in the crisis period the dynamics of the explanatory power of CHP and **CHP** part ways (@tbl-time-reg-US-CHP), seemingly reverting to the Keynesian paradigm observed pre-financialisation, further suggesting a return to legacy commodity specific dynamics. The recovery initiated in the crisis period for that of CHP is confirmed with a stronger increase, 27.2% vs. 22.5%, strongest for the metals sector, 37.3% vs. 23.7%, while the energy sector shows the opposite pattern: 11% vs. 16%. In contrast, after increasing stronger in the crisis, that for **CHP** falls overall post-crisis, 7.5% vs 14.7%, with the decrease strongest for the energy sector in a relative sense: 2.5% vs. 14.8%.

@@

```{r}
#| label: tbl-stats-regime-tests
#| tbl-cap: "This table shows mean returns and volatilities for the individual commodities as well as for two equally weighted portfolios of US and GB traded assets across the four periods. For each period, estimates are computed independently over the whole period (figure shown) as well as over phases of **backwardation** (b) and **contango** (c) with corresponding comparison results reported in brackets. The table further shows the results of t-tests of significance for the mean returns computed over the whole periods as well as t-tests of difference between these when computed over the two phases independently along with F-tests of difference for the corresponding variances. Values significant at the 1%, 5% and 10% level are marked with \\\\*\\\\*\\\\*, \\\\*\\\\* and \\\\* respectively. The estimate values per regime as well as the p-values for the corresponding difference tests are available in the attached [Internet Appendix](http://bautheac.duckdns.org:3838/bautheac/co-movement/). **Backwardation** and **contango** regimes as well as period definitions are discussed in \\@sec-methods while the results are discussed in \\@sec-results."

table_name <- "stats-individuals - regime difference tests"
columns_to_collapse_rows_for <- c("asset")
colnames <- c("asset", "estimate", "past", "financialisation", "crisis", "post-crisis")
align <- make_series_of_repeated_chars(list(l = 1L, r = 5L))
longtable <- TRUE
bold_columns <- 1L:6L 
coloured_columns <- 3L:6L

@@

```{r}
#| label: tbl-time-reg-US-CHP
#| tbl-cap: "This table shows the average time series $R^{2}$ for models where the returns series for the individual US traded commodities are independently regressed against relative change in their own commercial hedging pressure (CHP) series as well as against that in **CHP**, a market wide measure of CHP which calculation method is discussed in \\@sec-methods. Averages are computed independently across the whole cross-section of US traded commodity assets as well as across commodity sectors and sub-sectors for the four periods. Period definitions are discussed in \\@sec-methods while the results are discussed in \\@sec-results and the commodity assets taxonomy is detailed in \\@sec-taxonomy."

@@

```{r}
#| label: tbl-correlations-inner-periods
#| tbl-cap: "This table shows the average return pairwise correlation coefficients computed independently across the whole cross-section of individual commodity assets as well as within countries and commodity sectors and sub-sectors for the four periods. For each period, the results are computed independently over the entire period as well as over phases of **backwardation** and **contango**. **Backwardation** and **contango** regimes as well as period definitions are discussed in \\@sec-methods while the results are discussed in \\@sec-results and the commodity assets taxonomy is detailed in \\@sec-taxonomy."

@@

```{r}
#| label: tbl-correlations-cross-periods
#| tbl-cap: "This table shows the average return pairwise correlation coefficients between equally weighted US sub-sector portfolios (agriculturals: gains, livestock, softs; energy: gas, petroleum; metals: base, precious) independently computed for the four periods. For each period, the results are further computed independently over the entire period as well as over phases **backwardation** and **contango**. **Backwardation** and **contango** regimes as well as period definitions are discussed in \\@sec-methods while the results are discussed in \\@sec-results and the commodity assets taxonomy is detailed in \\@sec-taxonomy."

table_name <- "correlations - cross - US - periods - subsectors"
columns_to_collapse_rows_for <- c("regime")
colnames <- c("regime", periods)
align <- make_series_of_repeated_chars(list(l = 1L, r = 4L))
longtable <- TRUE
bold_columns <- 1L:5L
coloured_columns <- 2L:5L

@@

```{r}
#| label: tbl-correlations-inner-years
#| tbl-cap: "While \\@tbl-correlations-inner-periods shows results computed independently for the four periods, this table shows the results of the same analysis implemented for each year over the entire sample period. The results are discussed in \\@sec-results while \\@sec-taxonomy details the commodity assets taxonomy."

@@

```{r}
#| label: tbl-regressions-index
#| tbl-cap: "This table shows the average time series regression coefficients for models where the returns series for all the commodity assets are independently regressed against the return series for our market factor, an equally weighted portfolio of the US traded commodities. The results are computed independently across countries as well as commodity sectors and sub-sectors for the four periods. Period definitions along with the market factor construction method are discussed in \\@sec-methods while the results are discussed in \\@sec-results and the commodity assets taxonomy is detailed in \\@sec-taxonomy."

@@

```{r}
#| label: tbl-regressions-factors
#| tbl-cap: "This table shows the average time series regression $R^{2}$ for models where the returns series for all the commodity assets are independently regressed against the return series for our mimicking portfolios for risk factors in returns, including the market, CHP, open interest and term structure factors. The results are computed independently across countries and commodity sectors and sub-sectors for the four periods. A breakdown of the estimates computed independently over the whole periods as well as over phases of **backwardation** and **contango** is provided in the attached [Internet Appendix](http://bautheac.duckdns.org:3838/bautheac/co-movement/). Factor construction methods along with **backwardation** and **contango** regimes as well as period definitions are discussed in \\@sec-methods while the results are discussed in \\@sec-results and the commodity assets taxonomy is detailed in \\@sec-taxonomy."

table_name <- "regressions - all returns ~ factors - short"
columns_to_collapse_rows_for <- c("country", "sector", "subsector", "factor")
colnames <- c(sectors, "factor", periods)
align <- make_series_of_repeated_chars(list(l = 4L, r = 4L))
longtable <- TRUE
bold_columns <- 1L:8L
coloured_columns <- 5L:8L

@@

```{r}
#| label: tbl-assets-taxonomy
#| tbl-cap: "This table shows the detailed taxonomy for the thirty commodity assets considered in the study. The assets are classified by the country of residence of the exchange where they are traded at the time of writing as well as by sector and sub-sector according to the nature of the asset."
